Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-04-01 Paper/paper.tex
placeholder multiple observation figures
fig captions in S4

% #################
\begin{figure*}
\centering 
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
\caption{We show here (Log of) the ratio of the systematic and statistical
measurement uncertainties for the binary chirp mass over the NSBH parameter 
space. Each panel shows the same as a function of BH mass and spin. Across
each row, we can see the effect of increasing the signal strength (i.e. SNR)
with the NS's deformability fixed. Down each column, we can see 
the effect of the increasing NS's tidal deformability at fixed SNR. We also
show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
or $100\%$ (labelled ``0'').
% 
Generally, these two sources of uncertainty are different and for BBHs, the
statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
that for BHNS binaries, its not much different until we get to high SNRs
$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
measurement uncertainty for $\mchirp$ could become dominated by the exclusion
of tidal effects in filtering templates.
}
\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
\end{figure*}
%
\begin{figure*}[!t]
\centering    
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
with the difference that the quantity in question here is the symmetric 
mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
NS tidal effects on GW emission can become the dominant source of error. However,
for quieter signals with $\rho\leq 30$, we find that not including NS matter 
effects in parameter estimation templates has minimal effects on the measurement
of $\eta$. We remind the reader that the SNRs here are all single detector values.
}
\label{fig:TN_EtaBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
\begin{figure*}
\centering
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
\caption{These figures show the ratio of the systematic and statistical
measurement errors for BH spins, for all of our injected cases. The left
panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
tidal effects in templates introduces systematic effects that remain subdominant
to the statistical measurement uncertainties.
}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

@@

\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}

@@

$\lambdans$, as a percentage of the injected signal's $\lambdans$. In each panel, 

@@

panels with the same value of $\lambdans$, while each column contains panels 
with the same value of $\rho$. Two contours are drawn in each panel demarkating
that the statistical measurement uncertainty for $\lambda$ is equal to (or twice
of) its injected $\lambdans$ value, i.e. where we can place a factor of $1-2$ 
bound on $\lambdans$.

@@

\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\

@@

\caption{We demarkate here the regions of BH mass and spin plane where we can 

@@

Panels correspond to the injected SNR, with signal strength increasing
from left to right and top to bottom. Within each panel, $\lambdans$ measurement uncertainty
contours are shown as functions of BH mass and spin, with different colors corresponding to the
injected $\lambdans$. We note that this information is, in principle, contained in 
Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
the effect of NS's deformability itself on its measurement. As from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability of NS matter

@@

\pagebreak
\newpage
\FloatBarrier
\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
% 
% 
\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
% 
% 
Here we show a crude calculation which tells us that approximately $8$ (or $16$) observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$ (or $\rho=70$).

>>>>

% #################
\begin{figure*}
\centering 
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
\caption{We show here the ratio of systematic and statistical
measurement uncertainties for the binary chirp mass over the NSBH parameter 
space. Each panel shows the same as a function of BH mass and spin. Across
each row, we can see the effect of increasing the signal strength (i.e. SNR)
with the NS's deformability fixed. Down each column, we can see 
the effect of the increasing NS's tidal deformability at fixed SNR. We also
show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
or $100\%$ (labelled ``0'').
% 
Generally, these two sources of uncertainty are different and for BBHs, the
statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
that for BHNS binaries, its not much different until we get to high SNRs
$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
measurement uncertainty for $\mchirp$ could become dominated by the exclusion
of tidal effects in filtering templates.
}
\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
\end{figure*}
%
\begin{figure*}[!t]
\centering    
%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
\caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
with the difference that the quantity in question here is the symmetric 
mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
not including the effects of NS's tidal deformation on GW emission can become the dominant
source of error for astrophysical searches with Advanced LIGO. However,
for quieter signals with $\rho\leq 30$, it will have minimal effects on the measurement
of $\eta$. We remind the reader that the SNRs here are all single detector values.
}
\label{fig:TN_EtaBias_vs_Lambda_SNR}
\end{figure*}
% 
% 
\begin{figure*}
\centering
\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
\caption{This figure shows the ratio of the systematic and statistical
measurement errors for BH spins, with other attributed identical to 
Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
Similar to the case of mass parameters, we find that below
$\rho\approx 30$, ignoring tidal effects in templates introduces minor systematic effects,
which remain subdominant to the statistical measurement uncertainties.
}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

@@

@@

$\lambdans$, as a percentage of the injected signal's $\lambdans$ value. In each panel,

@@

panels with the same value of $\lambdans$, with $\rho$ increasing from left to right.
Each column contains panels with the same value of $\rho$, with $\lambdans$ 
increasing from top to bottom.
Contours are drawn in each panel demarkating regions where we can constrain the
$\lambdans$ parameter well (within a factor of two of the injected value).

@@

\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\

@@

\caption{In these figures we highlight the regions of BH mass and spin plane where we can 

@@

Panels correspond to different signal strength, with $\rho$ increasing from left to right.
Within each panel, $\lambdans$ measurement uncertainty contours are shown as functions of BH 
mass and spin, with line-types corresponding to different uncertainty levels, and
colors showing the injected $\lambdans$.
% We note that this information is, in principle, contained in 
% Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
% the effect of NS's deformability itself on its measurement. 
As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
of NS matter

@@

\pagebreak
\newpage
\FloatBarrier
\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
% 
% 
\begin{figure*}
\centering    
\includegraphics[width=1.\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
\caption{HELLO
}
\label{fig:TT_Lambda_vs_N_L800_CI90_0}
\end{figure*}
%
% 
\begin{figure*}
\centering    
\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L2000.pdf}
\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1500.pdf}\\
\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1000.pdf}
\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
\caption{HELLO
}
\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}
\end{figure*}
%
% 
% \begin{figure}
% \centering    
% \includegraphics[width=1.\columnwidth]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
% \caption{HELLO
% }
% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
% \end{figure}
%
% 
\begin{figure*}
\centering    
\includegraphics[width=1.\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
\includegraphics[width=1.\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
\caption{HELLO
}
\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
\end{figure*}
%

% 
\begin{figure}
\centering    
\includegraphics[width=1.\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N49_Log.pdf}
\caption{HELLO
}
\label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}
\end{figure}
%

% 
\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
% 
% 
Here we show a crude calculation which tells us that approximately $8$ (or $16$) observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$ (or $\rho=70$).
